Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-22 paper/example.tex
checkpointing ... mostly tweakage to examples section

The system receives an automation request that is forwarded to a third-party route planner (AI).
The route planner decides the flight path of the UAV based on its current position and the requested task.
The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes.
The waypoint manager is a repurposed legacy component that cannot be modified.

The expected behavior of the SW system, and the components implementing the system, are modeled with AGREE contract specifications.
The contracts constrain input with assumptions and state properties of output with guarantees.
AGREE performs model checking on this assume-guarantee system to hierarchically prove that the composite system obeys all contract obligations under all possible finite input streams.
The AGREE analysis, contract language, and the contracts for this example system are discussed in greater detail in \secref{sec:agree}.

The initial AGREE contracts for the system in \figref{fig:example}, and its implementing components, assume and guarantee the absence of any malicious, or unspecified, component behavior.
More precisely, the system assumes that all inputs are \emph{well-formed} and there is never more than one automation request pending at a time.
The meaning of well-formed is given by predicates for each data type.
For example, a waypoint is well-formed if it falls within bounds for latitude, longitude, and altitude.
The system guarantees that a start includes a waypoint, a start is within one cycle of an automation request and if not, then it persistently alerts, new waypoints coincide with air vehicle location updates, and all output are well-formed.

The components assume their inputs are well-formed, and they guarantee their outputs are well-formed.
The AI further guarantees it only responds to automation requests and always in the same cycle.
The legacy WM further guarantees a start from a response and always in the same cycle, start always comes with a waypoint, and further waypoints coincide with air vehicle location updates.
These initial specifications pass AGREE verification meaning that the contract composition of the components with the system satisfies all component input assumptions and system output guarantees.

A cyber-vulnerability analysis, however, identifies the potential of a supply chain attack through the AI route planner that is provided by a third-party vendor without source code.
The component is marked as \emph{untrusted} after the analysis, and its contract is modified by removing all guarantees on its outputs to reflect its untrusted status.
The AI output is now unconstrained in the assume-guarantee reasoning of AGREE and is able to generate any value on its output stream at anytime.

@@

The system implementation is cyber-hardened using BriefCASE, which automatically transforms the model by inserting high-assurance components in the form of a filter and a monitor as shown in \figref{fig:hardened}.
A filter enforces an invariant over each datum in the data stream by not forwarding input to its output if that input violates the filter invariant.
The auto-generated AGREE specification states that only well-formed inputs are passed to the output.
The system developer must provide this filtering policy, but it is usually based on the existing assumptions made by downstream components that consume the filter output.

@@

Preserving the input to output relationship of streams between the
AGREE contracts and CakeML lifts the AGREE contract verification
results to the actual deployed system.  If the contract model
verification succeeds, then the meaning of those results hold for the
deployed system.  These results, however, are only valid under
additional assumptions on the deployed system:
\begin{itemize}
\item first, contracts for non-synthesized components accurately model their deployed
counterparts;
\item second, an appropriate schedule exists to sequence component
  execution following the dependent data-flow in the design; and
\item third the communication fabric stitching components together works in harmony
  with the schedule.
\end{itemize}
\noindent Support and automation for these aspects of the design process are
discussed in other works.\footnote{KLS:need cites to HAMR, seL4, and Adventium work}

>>>>

The system receives an automation request that is forwarded to a
third-party route planner (AI).  The route planner decides the flight
path of the UAV based on its current position and the requested task.
The waypoint manager (WM) receives the mission command as a set of
waypoints from the planner and starts the UAV flying the mission,
issuing waypoints to the UAV flight controller as the UAV location
changes.  The waypoint manager is a legacy component that cannot be
modified.

The expected behavior of the SW system, and the components
implementing the system, are modeled with AGREE contract
specifications.  The contracts constrain input with assumptions and
state properties of output with guarantees.  AGREE performs model
checking on this assume-guarantee system to hierarchically prove that
SW obeys all contract obligations under all possible finite input
streams.  The AGREE analysis, contract language, and contracts for SW
are discussed in greater detail in \secref{sec:agree}.

The initial AGREE contracts for the system in \figref{fig:example},
and its implementing components, both assume and guarantee the absence
of any malicious, or unspecified, component behavior.  More precisely,
the system contract assumes that all inputs are \emph{well-formed} and
there is never more than one automation request pending at a time.
The meaning of well-formed is given by user-written predicates for
each data type.  For example, a waypoint is well-formed if it falls
within bounds for latitude, longitude, and altitude.  The guarantees
for SW ensure that
\begin{itemize}
\item a start includes a waypoint,\footnote{KLS: maybe describe the start port earlier.}
\item a start is within one cycle of an automation request and if not, then it persistently alerts,
\item new waypoints coincide with air vehicle location updates, and
\item all outputs are well-formed.
\end{itemize}

The contracts for the sub-components assume their inputs are
well-formed, and they guarantee their outputs are well-formed.  The AI
contract guarantees it only responds to automation requests and always
in the same cycle.  The legacy WM contract guarantees the following: a
start from a response and always in the same cycle; start always comes
with a waypoint; and, further waypoints coincide with air vehicle
location updates.  These initial specifications pass AGREE
verification meaning that the contract composition of the components
with the system satisfies all component input assumptions and system
output guarantees.

\subsection{Detecting and addressing vulnerabilities}
A cyber-vulnerability analysis, however, identifies the potential of a
supply chain attack through the AI route planner (provided by a
third-party vendor without source code).  BriefCASE marks the
component as \emph{untrusted} after the analysis, and the contract for
AI is modified by removing all guarantees on its outputs to reflect
its untrusted status.  The AI output is now unconstrained in the
assume-guarantee reasoning of AGREE and is able to generate any value
on its output stream at anytime.

@@

The system designer now uses BriefCASE to cyber-harden SW by inserting
high-assurance components in the form of a filter and a monitor, as
shown in \figref{fig:hardened}.  A filter enforces an invariant over
each datum in the data stream by not forwarding input to its output if
that input violates the filter invariant.  The auto-generated AGREE
specification states that only well-formed inputs are passed to the
output.  The system developer must further define the filtering
policy, but it is usually based on the existing assumptions made by
downstream components that consume the filter output.

@@

Preserving the input to output relationship of streams between the
AGREE contracts and CakeML lifts the AGREE contract verification
results to the actual deployed system.  If the contract model
verification succeeds, then the meaning of those results hold for the
deployed system.  These results, however, are only valid under
additional assumptions on the deployed system:
\begin{itemize}
\item contracts for non-synthesized components accurately model their deployed
counterparts;
\item an appropriate schedule exists to sequence component
  execution following the dependent data-flow in the design; and
\item the communication fabric stitching components together works in harmony
  with the schedule.
\end{itemize}
\noindent Support and automation for these aspects of the design process are
discussed in other works.\footnote{KLS:need cites to HAMR, seL4, and Adventium work}

### 2022-03-22 paper/verification-conditions.tex
checkpointing ... mostly tweakage to examples section

The syntax of AGREE is essentially that of quantifier-free first order
logic (FOL) supplemented with a few temporal operators. The terms
(\emph{e}) are the usual arithmetic expressions built from variables
($v$) and numeric and boolean literals ($c$), while formulas
(\emph{b}) are built using the standard logical connectives from
atomic formulas (\emph{a}) based on the usual comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]

Supplementing this basic syntax are is a conditional,
$\itelse{b}{(-)}{(-)}$, for both terms and formulas. There are also
temporal operators $\konst{pre}(-)$, delay $(-) \to (-)$, and
$\konst{Hist}(-)$.

>>>>

The syntax of AGREE is essentially that of quantifier-free first order
logic (FOL) supplemented with a few temporal operators. The terms
(\emph{e}) are arithmetic expressions built from variables ($v$) and
numeric and boolean literals ($c$), while formulas (\emph{b}) are
built using logical connectives from atomic formulas (\emph{a}) based
on the familiar comparison operators.
\[
\begin{array}{rcl}
e & ::= & v \mid c \mid e \;\set{+,*,/}\; e \\
a & ::= & e\; \set{=,<}\; e \\
b & ::= & v \mid c \mid a \mid \neg b
            \mid b \; \set{\land,\lor,\imp,\iff}\; b
\end{array}
\]

There is also a conditional, $\itelse{b}{(-)}{(-)}$, for both terms
and formulas. Lastly, there are temporal operators $\konst{pre}(-)$,
delay $(-) \to (-)$, and $\konst{Hist}(-)$.
